---
title: "Assignment 2 - Functions, iteration and debugging"
author: "Giulia Maria Petrilli"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
---
  
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

<!-- Do not forget to input your Github username in the YAML configuration up there --> 

***

```{r, include = T}
library(tidyverse)
library(dplyr)
library(unvotes)
library(lubridate)
library(dplyr)
```

<br>

***


### Task 1 - Fun with functions

a) Program a function `ultimate_answer()` that always returns the number 42 regardless of which input is provided, and show that it works providing three function calls that test different input types!

```{r}

# Running an if loop 
ultimate_answer <- function(x){
  if (is.numeric(x)) { # If x is numeric, print 42 
  return(42)
  } else {   # Print 42 in any other case 
    return(42)
  } 
}

# Checking if it works 

ultimate_answer(55)
ultimate_answer("Hello")
ultimate_answer(FALSE)

# This works but it takes long. Let's try with a for loop 

vector <- c(55, "Hello", FALSE)  # Creating a vector to test my loop
result <- c() # Store the results 
for (i in vector) {  # for every element  in vector 
  temp <- ultimate_answer(i)
  result <- c(result, temp)
}

result 

# Much better 


```

b) Write an R function called `color_guessing_game()` that allows the player to guess one of four colors (red, yellow, green, blue). The game will randomly select one of the colors and prompt the player to guess it. If the player's guess matches the selected color, they win, otherwise, they lose. *Hint:* You will need the function `readline()` to enable interactive communication between the game and the player.

```{r}

colours <- c("red", "yellow", "green", "blue") # Create colours vector 

color_guessing_game <- function() { 
  correct_answer <- sample(colours, 1) # sample a correct answer from colours 
  ANSWER <- readline("Guess what colour I am thinking about :)")
  if (!ANSWER %in% colours)
    stop("R does not know this colour!") # if the user inserts anything not in the vector,  error message. 
  if (ANSWER == correct_answer)
    cat("Yes! R's favourite colour")
  else
    cat("Wrong colour :/") # Else, print this 
}

if(interactive()) color_guessing_game()  # To make the game interactive 

```

<br>

***

### Task 2 - Functional programming in practice

The `unvotes` package (hosted both on [CRAN](https://cran.r-project.org/web/packages/unvotes/index.html) and [GitHub](https://github.com/dgrtwo/unvotes)) provides data on the voting history of countries in the [United Nations General Assembly](http://www.un.org/en/ga/), along with information such as date, description, and topics for each vote. The package contains three datasets: `un_votes`, providing the history of each country's votes, `un_roll_calls`, providing information on each roll call vote, and `un_roll_call_issues`, providing issue (topic) classifications of roll call votes. Check out the [database tutorial](https://github.com/dgrtwo/unvotes) if you want to make yourself familiar with the data structure. Then, work on the following tasks.

a) Calculate how often, on average, Germany agreed with the US on a vote in the UN general assembly since 1990. Votes where voting information for one of the countries is missing should be discarded.

```{r}

df <- # US and DE agreements and disagreements  
  (left_join(un_votes,un_roll_calls)) |>
  filter(vote != "abstain",  # Drop abstain
         country == "Germany" |  country == "United States", # Countries of interest 
         date >= "1990-01-01"  )  # Date range 


total_we_divide_by <- round(length(unique(df$rcid)), 5) # Duplicate removed, this is the total 


# Select the agreements only 

df <- df |> 
  group_by(rcid) |>  #Group by  roll call id
      filter(n() == 2) |> # We are only interested in roll calls with two entries for each country 
                          # Single roll calls do not allow for finding pairs between Germany and US agreements
      select(rcid, vote) |># Only selecting the columns of interest 
      filter(duplicated(vote)) # To avoid repetition of the rcid and summarize by agreements 


(nrow(df) / (total_we_divide_by)) # Agreements / total 



```

<br>

b) Now, create a function, `votes_agreement_calculator()`, that takes country identifiers as well as a `year_min` argument as inputs and that returns the share of agreement in voting between any two specified countries as numeric value, for a time period specified with year >= `year_min`. The function should take the necessary data frames directly from the `unvotes` package. Then, test the function by computing the agreement rates for (a) the United States and Russia for votes cast in 2000 and later and (b) France and the UK for votes cast in 2000 and later!

```{r}

votes_agreement_calculator <- function(code,  year_min){
  df <-  (left_join(un_votes,un_roll_calls, join_by(rcid))) |>
     mutate(year = lubridate::year(date)) |> # only want YYYY so it recognizes year_min
  filter(vote != "abstain",
         # Abstain does not mean agreement. Even if two countries are both abstaining, they could be abstaining          for different reasons
         country_code %in% code,
         year >= year_min) |>
  pivot_wider(id_cols = c(rcid, date, session), names_from = country , values_from = vote) |> # every country has a column
    drop_na()
  
    last_two_columns <- df[, (ncol(df) - 1):ncol(df)] # Select the last two columns, as I know the two countries will be there. Not the most robust but I am only working on this df for this excercise 
    matches <- as.numeric(last_two_columns[,1] == last_two_columns[,2])
    round(mean(matches), 3) * 100
  
}

US_RU <- votes_agreement_calculator(c("US", "RU"), 2000)
FRA_GB <- votes_agreement_calculator(c("FR", "GB"), 2000)
```


<br>

c) Using `purrr` functionality, find out which three countries on average agreed with the US the most from the year 2000 on!

```{r, eval = TRUE}

  
un_vote_wide <- un_votes |> 
  inner_join(un_roll_calls, join_by(rcid)) |> 
  mutate(country = str_replace_all(country, " ", "_")) |>   # replace string 
  pivot_wider(id_cols = c(rcid, date, session), names_from = country, values_from = vote) # all countries have columns 

US_agree <- function(country_1, year_min) {
  un_vote_wide <- un_votes |> 
    inner_join(un_roll_calls, join_by(rcid)) |> 
    mutate(country = str_replace_all(country, " ", "_")) |> 
    pivot_wider(id_cols = c(rcid, date, session), names_from = country, values_from = vote) 

  US_agree_avg <- un_vote_wide |>   # within function, define we need the US 
    select(rcid, session, date, all_of("United_States"), all_of({{country_1}})) |> 
    mutate(date = as.Date(date)) |>    # change date 
    filter(year(date) >= {{year_min}}) |> 
    drop_na() |> 
    mutate(agreement = ifelse(!!sym("United_States") == !!sym(country_1), 1, 0)) |> 
    summarise(average_agreement = mean(agreement)*100)
  
  return(US_agree_avg)
}

country_columns <- names(un_vote_wide) |> 
  keep(~ str_detect(., "^[A-Z]") & .x != "United_States")  

US_agree_avg_map <- map(country_columns, ~ US_agree(.x, 2000)) # function to each col with map 

result <- data.frame(unlist(US_agree_avg_map), country_columns) |>   # transform list into df to arrage best 
  setNames(c("avg_agree", "country"))

result |>    # Top three countries that most agreed w the us 
  arrange(desc(avg_agree)) |>  # descending order 
  head(3)  # to select 3 



```

<br>

***

### Task 3 - Debugging code

The following code snippet contains various bugs. Flag them and suggest a fix by adding a comment in the respective line. Example:

```{r, eval = FALSE}
library(Tidyverse) # BUG: typo in library(Tidyverse); should be library(tidyverse) instead
```

```{r, eval = FALSE}
# load packages
library(tidyverse) # This is correct 
library(countrycode) #
library(Unvotes) # unvotes, all lower case 

# get continents data   

continents <- countrycode::codelist %>% # 2 semicolon 
  select(continent, iso2c) %>%
  rename(country_code = iso2c)  # just 1 = 

un_votes <- left_join(x = un_votes, y = continents, by = "country_code", relationship = "many-to-many") 

'''A country_code can be connected to multiple rows in continents, and a continent can be connected to multiple rows in, creating a many-to-many relationship between them.'''


# get data on European UN votes
eu_un_votes <- left_join(un_votes, un_roll_calls, by = "rcid") %>%
  left_join(., un_roll_call_issues, by = "rcid") %>%  # BUG: Unnecessary . ; FIX: Remove the .
  filter(continent == "Europe", 
         date >= "1991-01-01" & date <= "2021-12-31") %>% 
  drop_na(short_name) 

# Fixed 
# encode Europe's voting behavior numerically
eu_un_votes <- eu_un_votes %>% 
  mutate(yes_vt = if_else(vote == "yes", 1, 0)) # No comma at the end 
         no_vt = if_else(vote == "no", 1, 0),
         abstention = if_else(vote == "abstain", 1, 0))
# list top 10 most disagreed with UN resolutions
hated_un <- eu_un_votes %>%
  group_by(unres)
  summarise(sh_disagree_with = sum(no_vt,na.rm = T) / sum(c(yes_vt,no_vt, abstention), na.rm = T)) %>% 
  arrange(desc(sh_disagree_with)) %>% 
  head(10)
hated_un
``` 


